---
title: "AI: Aleutian Islands US survey data processing summary"
author: "fishglob, Aurore A. Maureaud, Juliano Palacios Abrantes, Zoë Kitchel, Dan Forrest, & Michelle Stuart"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: true
  # html_notebook:
  #   fig_width: 6
  #   fig_height: 4
  #   toc: yes
  #   toc_depth: 3
  #   code_folding: hide
  #   toc_float:
  #     collapsed: no
  # html_document:
  #   toc: yes
  #   toc_depth: '3'
  #   df_print: paged
---

```{r setup, include=FALSE, results='hide'}

library(readxl) # To load extra datasets
library(tidyverse) # for data wrangling
library(janitor) # for cleaning names
library(lubridate) # for fixing dates
library(rnaturalearth) # for removing points from land
library(sf) # for removing points from land
library(sp) # for removing points from land
library(here) # for easy work around on multiple computers
library(taxize) # for getting correct species names
library(magrittr) # for names wrangling
library(googledrive) # for interaction with the drive
library(knitr) # for nice tables
library(kableExtra)
library(plotly) # for interactive plots
library(leaflet) # for interactive map
library(mapdata)
library(data.table)

# Personal functions
source(here::here("functions/write_clean_data.R"))
source(here("functions/clean_taxa.R"))

fishglob_data_columns <- read_excel(here("standard_formats/fishglob_data_columns.xlsx"))

```


## General info

This document presents the summary of the Aleutian Island bottom trawl survey provided by [Stan Kotwicki](stan.kotwicki@noaa.gov) and [Jim Thorson](james.thorson@noaa.gov) It contains data from 1983-1997 (triennial) and 2000-2020 (biennial; 2008 cancelled).


## Data cleaning in R

```{r cleaning_code, code = readLines(here("./cleaning_codes/get_ai.R")), eval = FALSE}

```


```{r load_data, echo=F, results='hide', include=FALSE, eval=T}

### Survey data

# drive_download(file = "AI_clean.csv",
#   overwrite = TRUE)
# 
# survey <- read.csv(here("./summary/AI_clean.csv"))

load(here("outputs/Cleaned_data/AI_clean.RData"))
survey <- data
rm(data)

### Map data
World_map <- rnaturalearth::ne_countries(scale = 'medium', returnclass = c("sf")) %>%
  clean_names() %>% 
  select(iso_a3,sovereignt,sov_a3,name,name_long,region_wb,continent,geometry) %>% 
  # st_transform(crs = "+proj=eck4") %>%
  st_transform(crs = 4326) %>% # 4326
  mutate(iso_a3 = ifelse(sovereignt == "Norway","NOR",
                         ifelse(name == "France","FRA",
                                iso_a3)
                         )
         )

```

\clearpage

## 1. Overview of the survey data table

```{r head_survey, eval = T, echo = F}
kable(survey[1:5,1:6], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,7:15], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,16:23], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,24:30], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,31:35], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,36:42], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

```

\clearpage

## 2. Summary of sampling intensity

Number of hauls per year performed during the survey after data processing.

```{r hauls_per_year, eval = T, echo = F, fig.width = 5, warning=F, message=F}

survey %>% 
  # head()
  group_by(year,haul_id) %>% 
  summarise(n =  n()) %>% 
  group_by(year) %>% 
  summarise(n_hauls = n()) %>% 
  ggplot()  +
  geom_line(aes(x = year,y = n_hauls)) +
  geom_point(aes(x = year,y = n_hauls)) +
  ylab("Number of hauls") + xlab("Year") + theme_bw()

```


\clearpage

## 3. Summary of sampling variables from the survey

Here we show the yearly total and average of the following variables reported in the survey data:

- *area_swept*, swept area by the bottom trawl gear ${km^2}$
- *depth*, sampling depth in ${m}$
- *haul_dur*, haul sampling duration ${hour}$
- *number of marine fish taxa*, taxa were cleaned following the last version of taxonomy from the World Register of Marine Species (https://www.marinespecies.org/, October 2021)


```{r summary_tech_vars, eval = T, echo = F, fig.width=10, fig.height = 6, message =  F,warning =  F}

var_plot <-  survey %>%
  group_by(year, haul_id, haul_dur, area_swept, depth) %>%
  summarize(nbr_taxa = length(accepted_name)) %>% 
  gather("var","val",3:6) %>% 
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = val),
    outlier.shape = NA,
    size=0.5
  )  +
  # geom_line(
  #   aes(
  #     x = year,
  #     y = val
  #   )
  # )  +
  ylab("")  + xlab("Year") +
  facet_wrap(~var, 
             scales = "free_y",
             ncol = 2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
  
var_plot

```

\clearpage

## 4. Summary of biological variables

Here we display the yearly total and average across hauls of the following variables recorded in the data:

- *num_cpua*, number of individuals (abundance) in $\frac{individuals}{km^2}$
- *num_cpue*, number of individuals (abundance) in $\frac{individuals}{h}$
- *num*, number of individuals (abundance)
- *wgt_cpua*, weight in $\frac{kg}{km^2}$
- *wgt_cpue*, weight in $\frac{kg}{h}$
- *wgt*, weight in ${kg}$

```{r summary_var_plot, eval = T, echo = F, message =  F,warning =  F}

var_plot <-  survey %>%
  group_by(year) %>% 
  summarise_at(vars(num:wgt_cpua),
               funs(sum,mean),na.rm=T) %>% 
  # head()
  gather("var","val",2:13) %>% 
  # head()
  ggplot() +
  geom_point(
    aes(
      x = year,
      y = val
    ),
    size=0.5
  )  +
  geom_line(
    aes(
      x = year,
      y = val
    ),
    size=0.5
  )  +
  ylab("")  + xlab("Year") +
  facet_wrap(~var, 
             scales = "free_y",
             ncol = 3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
  

var_plot

```


\clearpage

## 5. Extreme values

Here we show a yearly total distribution of the biomass data to visualize outliers:

- *num_cpue*, number of individuals (abundance) in $\frac{individuals}{km^2}$
- *wgt_cpue*, weight in $\frac{kg}{km^2}$

```{r extreme_biomass, eval = T, echo = F, message =  F,warning =  F}

if(!is.na(mean(survey$num_cpua, na.rm=T)) & !is.na(mean(survey$wgt_cpua, na.rm=T))){
  var_plot <-  survey %>%
  group_by(year, haul_id) %>% 
  summarize(Weight = sum(wgt_cpua), Abundance = sum(num_cpua)) %>% 
  gather("var","val",3:4) %>% 
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = val
    ),
    size.outlier = 0.5,
    size = 0.5
  )  +
  facet_wrap(~var, scales = "free_y",ncol = 2) +
  ylab("")  + xlab("Year") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))
}

if(!is.na(mean(survey$num_cpua, na.rm=T)) & is.na(mean(survey$wgt_cpua, na.rm=T))){
var_plot <-  survey %>%
  group_by(year, haul_id) %>% 
  summarize(Abundance = sum(num_cpua)) %>% 
  # head()
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = Abundance
    ),
    size.outlier = 0.5,
    size = 0.5
  )  +
  ylab("Abundance")  + xlab("Year") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))
} 

if(is.na(mean(survey$num_cpua, na.rm=T)) & !is.na(mean(survey$wgt_cpua, na.rm=T))){
var_plot <-  survey %>%
  group_by(year, haul_id) %>% 
  summarize(Weight = sum(wgt_cpua)) %>% 
  # head()
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = Weight
    ),
    size.outlier = 0.5,
    size = 0.5
  )  +
  ylab("Weight")  + xlab("Year") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))
} 
var_plot

```


\clearpage

## 6. Summary of variables against swept area

Here we show the total abundance and number of taxa relationships with the area swept:

- *nbr_taxa*, number of marine fish taxa after taxonomic data cleaning
- *num_cpua*, number of individuals (abundance) in $\frac{individuals}{km^2}$
- *wgt_cpua*, weight in $\frac{kg}{km^2}$



```{r summary_var_swept, eval = T, echo = F, message =  F,warning =  F}

if(!is.na(mean(survey$num_cpua, na.rm=T)) & !is.na(mean(survey$wgt_cpua, na.rm=T))){
  var_plot <-  survey %>%
  group_by(haul_id, haul_dur, area_swept) %>%
  summarize(Number_Taxa = length(accepted_name), Abundance = sum(num_cpua),Weight = sum(wgt_cpua)) %>% 
  gather("var","val",4:6) %>% 
  # head()
  ggplot() +
  geom_point(aes(x = area_swept,y = val),size=0.5)  +
  ylab("")  + xlab("Swept Area") +
  facet_wrap(~var, scales = "free_y", ncol = 3) +
  theme_bw()
}

if(!is.na(mean(survey$num_cpue, na.rm=T)) & mean(survey$wgt_cpue, na.rm=T)){
  var_plot <-  survey %>%
  group_by(haul_id, haul_dur, area_swept) %>%
  summarize(Number_Taxa = length(accepted_name), Abundance = sum(num_cpue)) %>% 
  gather("var","val",4:5) %>% 
  # head()
  ggplot() +
  geom_point(aes(x = area_swept, y = val),size=0.5)  +
  ylab("")  + xlab("Swept Area") +
  facet_wrap(~var,  scales = "free_y", ncol = 2) +
  theme_bw()
}

if(is.na(mean(survey$num_cpua, na.rm=T)) & !is.na(mean(survey$wgt_cpua, na.rm=T))){
  var_plot <-  survey %>%
  group_by(haul_id, haul_dur, area_swept) %>%
  summarize(Number_Taxa = length(accepted_name), Weight = sum(wgt_cpua)) %>% 
  gather("var","val",4:5) %>% 
  # head()
  ggplot() +
  geom_point(aes(x = area_swept, y = val),size=0.5)  +
  ylab("")  + xlab("Swept Area") +
  facet_wrap(~var,  scales = "free_y", ncol = 2) +
  theme_bw()
}

var_plot
```

\clearpage

## 7. Abundance or Weight trends of the six most abundant species

```{r abundant_spp, eval=T, echo=F, message=F, warning=F}

if(!is.na(mean(survey$num_cpua, na.rm=T))){
spp <- survey %>% 
  group_by(year, accepted_name) %>% 
  summarize(wgt = sum(wgt_cpua), nbr_years = length(year)) %>% 
  filter(nbr_years>10) %>% 
  group_by(accepted_name) %>% 
  summarize(wgt = median(wgt)) %>% 
  arrange(desc(wgt)) %>% 
  top_n(n=6) %>% 
  pull(accepted_name)

spp_plot <- survey %>% 
  filter(accepted_name %in% spp) %>% 
  group_by(year, accepted_name) %>%
  summarize(wgt = sum(wgt_cpua, na.rm=T)) %>% 
  ggplot() +
  geom_point( aes(x = year, y = wgt), size=0.5 ) +
  geom_line(aes(x = year,y = wgt), size=0.5) +
  facet_wrap(~accepted_name,  scales = "free_y",ncol = 2) +
  theme_bw() +
  ylab("Species Weight (kg)") + xlab("Year")
}

if(is.na(mean(survey$wgt_cpua, na.rm=T))){
  spp <- survey %>% 
  group_by(year, accepted_name) %>% 
  summarize(num = sum(num_cpua), nbr_years = length(year)) %>% 
  filter(nbr_years>10) %>% 
  group_by(accepted_name) %>% 
  summarize(num = median(num)) %>% 
  arrange(desc(num)) %>% 
  top_n(n=6) %>% 
  pull(accepted_name)

spp_plot <- survey %>% 
  filter(accepted_name %in% spp) %>% 
  group_by(year, accepted_name) %>%
  summarize(num = sum(num_cpua, na.rm=T)) %>% 
  ggplot() +
  geom_point( aes(x = year, y = num), size=0.5 ) +
  geom_line(aes(x = year,y = num), size=0.5) +
  facet_wrap(~accepted_name,  scales = "free_y",ncol = 2) +
  theme_bw() +
  ylab("Species Abundance") + xlab("Year")
}

spp_plot
```

\clearpage

## 8. Distribution mapping

Map of the sampling distribution in space. Note that we only show one year per coordinate.

```{r fixed_point_map, eval = T, echo = F, fig.width=10, fig.height= 5, message =  F,warning =  F}

world <- data.table(map_data('world'))
north_america <- world[long >= -200 & long <= -130 & lat >= 50 & lat <= 80]
#if positive, subtract 360
survey <- survey %>%
  mutate(longitude_s = ifelse(longitude > 0,(longitude-360),(longitude)))
# Fixed map
survey %>% 
  select(longitude_s,latitude) %>% 
  distinct() %>% 
  ggplot() +
  geom_polygon(data = north_america, 
                                 aes(x=long, y = lat, group = group), 
                                 fill = "white", 
                                 color="black") +
    coord_fixed(xlim = c(-190, -160),  ylim = c(50, 60), ratio = 1.3) +
  geom_point(
    aes(
      x = longitude_s,
      y = latitude),
    alpha = 0.5
  ) +
  theme_bw() +
  theme(legend.position = "right")

```

\clearpage

## 9. Taxonomic flagging

This species flagging method was adapted from https://github.com/pinskylab/OceanAdapt/blob/master/R/add-spp-to-taxonomy.Rmd#L33

Visualization of flagged taxa

```{r, echo=FALSE, out.width = '80%'}
knitr::include_graphics(here::here("outputs", "Flags","taxonomic_flagging", paste0(survey$survey[1],"_taxonomic_flagging.png")))
```

Statistics related to the taxonomic flagging outputs

```{r, echo=FALSE}
df <- read.csv(here::here("outputs", "Flags","taxonomic_flagging", paste0(survey$survey[1],'_stats.csv')))
knitr::kable(df, col.names = NULL)
```

\clearpage 

## 10. Spatio-temporal standardization

### a. Standardization method 1

This standardization method was adapted from https://github.com/zoekitchel/trawl_spatial_turnover/blob/master/data_prep_code/species/explore_NorthSea_trimming.Rmd
\
It was run for hex resolution 7 and 8.

Plot of number of cells x years with overlaid flagging options

```{r, echo=FALSE, out.width = '80%'}
knitr::include_graphics(here::here("outputs", "Flags","trimming_method1", "hex_res7", paste0(survey$survey[1],"_hex_res_7_plot.png")))
```
```{r, echo=FALSE, out.width = '80%'}
knitr::include_graphics(here::here("outputs", "Flags","trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_plot.png")))
```

Map of hauls retained and removed per flagging method and threshold

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("outputs", "Flags","trimming_method1", "hex_res7",  paste0(survey$survey[1],"_hex_res_7_map_per_haul.png")))
```

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("outputs", "Flags", "trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_map_per_haul.png")))
```


Map of numbers of years removed per grid cell and flagging method/threshold

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("outputs", "Flags","trimming_method1", "hex_res7", paste0(survey$survey[1],"_hex_res_7_map_per_grid_nyears.png")))
```


```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("outputs", "Flags","trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_map_per_grid_nyears.png")))
```


### b. Standardization method 2

This standardization method was adapted from BioTIME code from https://github.com/Wubing-Xu/Range_size_winners_losers 
   
Map of hauls retained and removed

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("outputs", "Flags","trimming_method2", 
                                   paste0(survey$survey[1],"_map_per_haul.png")))
```

### c. Standardization summary

Statistics of hauls removed for each standardization method

```{r, echo=FALSE}
met1_7 <- read.csv(here::here("outputs", "Flags","trimming_method1", "hex_res7", paste0(survey$survey[1],"_hex_res_7_stats_hauls.csv")))
met1_8 <- read.csv(here::here("outputs", "Flags","trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_stats_hauls.csv")))
met2 <- read.csv(here::here("outputs", "Flags", "trimming_method2", 
                            paste0(survey$survey[1],"_stats_hauls.csv")))
knitr::kable(cbind(met1_7, met1_8[,2:3], met2[,2]), 
             col.names = c("summary", "grid cell 7, 0% threshold", "grid cell 7, 2% threshold", 
                           "grid cell 8, 0% threshold", "grid cell 8, 2% threshold",
                           "method 2 (biotime)")) %>% 
  kable_styling(latex_options = "scale_down",
                position = "float_left")
```

