---
title: "PT-IBTS: Portuguese Bottom Trawl survey data processing summary"
author: "fishglob, Aurore A. Maureaud & Juliano Palacios Abrantes & P.D van Denderen"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: true
---
---
title: "PT-IBTS: Portuguese Bottom Trawl survey data processing summary"
author: "fishglob, Aurore A. Maureaud & Juliano Palacios Abrantes & P.D van Denderen"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: true
---

```{r setup, include=FALSE, results='hide'}

library(readxl) # To load extra datasets
library(tidyverse) # for data wrangling
library(janitor) # for cleaning names
library(lubridate) # for fixing dates
library(rnaturalearth) # for removing points from land
library(sf) # for removing points from land
library(sp) # for removing points from land
library(here) # for easy work around on multiple computers
library(taxize) # for getting correct species names
library(magrittr) # for names wrangling
library(googledrive) # for interaction with the drive
library(knitr) # for nice tables
library(kableExtra)
library(plotly) # for interactive plots
library(leaflet) # for interactive map

# Personal functions
source(here("./functions/write_clean_data.r"))
source(here("./functions/clean_taxa.r"))

fishglob_data_columns <- read_excel(here("standard_formats/fishglob_data_columns.xlsx"))

```


## General info

This document presents the cleaning code and summary of the Portuguese bottom trawl survey provided by DATRAS, ICES. It contains data from **2002** and up to **2018**.


## Data cleaning in R


```{r cleaning_code, code = readLines(here("./cleaning.codes/get.datras.R")), eval=FALSE}
```


```{r load_data, echo=F, results='hide', include=FALSE, eval=T}

### Survey data
x <- drive_find(pattern = "DATRAS_v3_clean.csv")

drive_download(file = x[1,],
  overwrite = TRUE)

survey <- read.csv(here("./summary/DATRAS_v3_clean.csv")) %>% 
  filter(survey == "PT-IBTS")

### Map data
World_map <- rnaturalearth::ne_countries(scale = 'medium', returnclass = c("sf")) %>%
  clean_names() %>% 
  select(iso_a3,sovereignt,sov_a3,name,name_long,region_wb,continent,geometry) %>% 
  # st_transform(crs = "+proj=eck4") %>%
  st_transform(crs = 4326) %>% # 4326
  mutate(iso_a3 = ifelse(sovereignt == "Norway","NOR",
                         ifelse(name == "France","FRA",
                                iso_a3)
                         )
         )

```
\clearpage

## 1. Overview of the survey data table

```{r head_survey, eval = T, echo = F}
kable(survey[1:5,1:8], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,9:15], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,16:21], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,22:27], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,28:32], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))

kable(survey[1:5,33:39], format = "latex", booktabs = T) %>%
kable_styling(latex_options = c("striped","HOLD_position"))


```


\clearpage

## 2. Summary of sampling intensity

Number of hauls per year performed during the survey after data processing.

```{r hauls_per_year, eval = T, echo = F, fig.width = 5, warning=F, message=F}

survey %>% 
  # head()
  group_by(year,haul_id) %>% 
  summarise(n =  n()) %>% 
  group_by(year) %>% 
  summarise(n_hauls = n()) %>% 
  ggplot()  +
  geom_line(aes(x = year,y = n_hauls)) +
  geom_point(aes(x = year,y = n_hauls)) +
  ylab("Number of hauls") + xlab("Year") + theme_bw()

```



\clearpage

## 3. Summary of sampling variables from the survey

Here we show the yearly total and average of the following variables reported in the survey data:

- *area_swept*, swept area by the bottom trawl gear ${km^2}$
- *depth*, sampling depth in ${m}$
- *haul_dur*, haul sampling duration ${hours}$
- *number of marine fish taxa*, taxa were cleaned following the last version of taxonomy from the World Register of Marine Species (https://www.marinespecies.org/, October 2021)


```{r summary_tech_vars, eval = T, echo = F, fig.width=10, fig.height = 6, message =  F,warning =  F}

var_plot <-  survey %>%
  group_by(year, haul_id, haul_dur, area_swept, depth) %>%
  summarize(nbr_taxa = length(accepted_name)) %>% 
  gather("var","val",3:6) %>% 
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = val),
    outlier.shape = NA,
    size=0.5
  )  +
  # geom_line(
  #   aes(
  #     x = year,
  #     y = val
  #   )
  # )  +
  ylab("")  + xlab("Year") +
  facet_wrap(~var, 
             scales = "free_y",
             ncol = 2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
  
var_plot

```

\clearpage

## 4. Summary of biological variables

Here we display the yearly total and average across hauls of the following variables recorded in the data:

- *num_cpue*, number of individuals (abundance) in $\frac{individuals}{km^2}$
- *num_h*, number of individuals (abundance) in $\frac{individuals}{h}$
- *num*, number of individuals (abundance)
- *wgt_cpue*, weight in $\frac{kg}{km^2}$
- *wgt_h*, weight in $\frac{kg}{h}$
- *wgt*, weight in ${kg}$

```{r summary_var_plot, eval = T, echo = F, message =  F,warning =  F}

var_plot <-  survey %>%
  group_by(year) %>% 
  summarise_at(vars(num:wgt_cpue),
               funs(sum,mean),na.rm=T) %>% 
  # head()
  gather("var","val",2:13) %>% 
  # head()
  ggplot() +
  geom_point(
    aes(
      x = year,
      y = val
    ),
    size=0.5
  )  +
  geom_line(
    aes(
      x = year,
      y = val
    ),
    size=0.5
  )  +
  ylab("")  + xlab("Year") +
  facet_wrap(~var, 
             scales = "free_y",
             ncol = 3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
  

var_plot

```


\clearpage

## 5. Extreme values

Here we show a yearly total distribution of the biomass data to visualize outliers:

- *wgt*, total weight in ${kg}$ per haul and year per haul and year, if available in the survey data
- *num*, total number of individuals, if available in the survey data

```{r extreme_biomass, eval = T, echo = F, message =  F,warning =  F}

if(!is.na(mean(survey$num_cpue, na.rm=T)) & !is.na(mean(survey$wgt_cpue, na.rm=T))){
  var_plot <-  survey %>%
  group_by(year, haul_id) %>% 
  summarize(Weight = sum(wgt_cpue), Abundance = sum(num_cpue)) %>% 
  gather("var","val",3:4) %>% 
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = val
    ),
    size.outlier = 0.5,
    size = 0.5
  )  +
  facet_wrap(~var, scales = "free_y",ncol = 2) +
  ylab("")  + xlab("Year") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))
}

if(!is.na(mean(survey$num_cpue, na.rm=T)) & is.na(mean(survey$wgt_cpue, na.rm=T))){
var_plot <-  survey %>%
  group_by(year, haul_id) %>% 
  summarize(Abundance = sum(num_cpue)) %>% 
  # head()
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = Abundance
    ),
    size.outlier = 0.5,
    size = 0.5
  )  +
  ylab("Abundance")  + xlab("Year") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))
} 

if(is.na(mean(survey$num_cpue, na.rm=T)) & !is.na(mean(survey$wgt_cpue, na.rm=T))){
var_plot <-  survey %>%
  group_by(year, haul_id) %>% 
  summarize(Weight = sum(wgt_cpue)) %>% 
  # head()
  ggplot() +
  geom_boxplot(
    aes(
      x = as.factor(year),
      y = Weight
    ),
    size.outlier = 0.5,
    size = 0.5
  )  +
  ylab("Weight")  + xlab("Year") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))
} 
var_plot
rm(var_plot)
```


\clearpage

## 6. Summary of variables against swept area

Here we show the total abundance and number of taxa relationships with the area swept:

- *nbr_taxa*, number of marine fish taxa after taxonomic data cleaning
- *num*, number of individuals, if available in the survey data
- *wgt*, weight in ${kg}$, if available in the survey data


```{r summary_var_swept, eval = T, echo = F, message =  F,warning =  F}

if(!is.na(mean(survey$num, na.rm=T)) & !is.na(mean(survey$wgt, na.rm=T))){
  var_plot <-  survey %>%
  group_by(haul_id, haul_dur, area_swept) %>%
  summarize(Number_Taxa = length(accepted_name), Abundance = sum(num),Weight = sum(wgt)) %>% 
  gather("var","val",4:6) %>% 
  # head()
  ggplot() +
  geom_point(aes(x = area_swept,y = val),size=0.5)  +
  ylab("")  + xlab("Swept Area") +
  facet_wrap(~var, scales = "free_y", ncol = 3) +
  theme_bw()
}

if(!is.na(mean(survey$num, na.rm=T)) & is.na(mean(survey$wgt, na.rm=T))){
  var_plot <-  survey %>%
  group_by(haul_id, haul_dur, area_swept) %>%
  summarize(Number_Taxa = length(accepted_name), Abundance = sum(num)) %>% 
  gather("var","val",4:5) %>% 
  # head()
  ggplot() +
  geom_point(aes(x = area_swept, y = val),size=0.5)  +
  ylab("")  + xlab("Swept Area") +
  facet_wrap(~var,  scales = "free_y", ncol = 2) +
  theme_bw()
}

if(is.na(mean(survey$num, na.rm=T)) & !is.na(mean(survey$wgt, na.rm=T))){
  var_plot <-  survey %>%
  group_by(haul_id, haul_dur, area_swept) %>%
  summarize(Number_Taxa = length(accepted_name), Weight = sum(wgt)) %>% 
  gather("var","val",4:5) %>% 
  # head()
  ggplot() +
  geom_point(aes(x = area_swept, y = val),size=0.5)  +
  ylab("")  + xlab("Swept Area") +
  facet_wrap(~var,  scales = "free_y", ncol = 2) +
  theme_bw()
}

var_plot

```

\clearpage

## 7. Abundance or Weight trends of the six most abundant species

```{r abundant_spp, eval=T, echo=F, message=F, warning=F}

if(!is.na(mean(survey$wgt_cpue, na.rm=T))){
spp <- survey %>% 
  group_by(year, accepted_name) %>% 
  summarize(wgt = sum(wgt_cpue), nbr_years = length(year)) %>% 
  filter(nbr_years>10) %>% 
  group_by(accepted_name) %>% 
  summarize(wgt = median(wgt)) %>% 
  arrange(desc(wgt)) %>% 
  top_n(n=6) %>% 
  pull(accepted_name)

spp_plot <- survey %>% 
  filter(accepted_name %in% spp) %>% 
  group_by(year, accepted_name) %>%
  summarize(wgt = sum(wgt_cpue, na.rm=T)) %>% 
  ggplot() +
  geom_point( aes(x = year, y = wgt), size=0.5 ) +
  geom_line(aes(x = year,y = wgt), size=0.5) +
  facet_wrap(~accepted_name,  scales = "free_y",ncol = 2) +
  theme_bw() +
  ylab("Species Weight (kg)") + xlab("Year")
}

if(is.na(mean(survey$wgt_cpue, na.rm=T))){
  spp <- survey %>% 
  group_by(year, accepted_name) %>% 
  summarize(num = sum(num_cpue), nbr_years = length(year)) %>% 
  filter(nbr_years>10) %>% 
  group_by(accepted_name) %>% 
  summarize(num = median(num)) %>% 
  arrange(desc(num)) %>% 
  top_n(n=6) %>% 
  pull(accepted_name)

spp_plot <- survey %>% 
  filter(accepted_name %in% spp) %>% 
  group_by(year, accepted_name) %>%
  summarize(num = sum(num_cpue, na.rm=T)) %>% 
  ggplot() +
  geom_point( aes(x = year, y = num), size=0.5 ) +
  geom_line(aes(x = year,y = num), size=0.5) +
  facet_wrap(~accepted_name,  scales = "free_y",ncol = 2) +
  theme_bw() +
  ylab("Species Abundance") + xlab("Year")
}

spp_plot

```

\clearpage

## 8. Distribution mapping

Map of the sampling distribution in space. Note that we only show one year per coordinate.

```{r fixed_point_map, eval = T, echo = F, fig.width=10, fig.height= 5, message =  F,warning =  F}

# Fixed map
survey %>% 
  select(longitude,latitude) %>% 
  distinct() %>% 
  ggplot() +
  geom_point(
    aes(
      x = longitude,
      y = latitude),
    alpha = 0.5
  ) +
  geom_sf(data = World_map, aes(), alpha = 0.5) +
  coord_sf(xlim = c(min(survey$longitude, na.rm=T), max(survey$longitude, na.rm=T)),
           ylim = c(min(survey$latitude, na.rm=T), max(survey$latitude, na.rm=T))
  ) +
  theme_bw() +
  theme(legend.position = "right")

```



## 9. Taxonomic flagging

This species flagging method was adapted from https://github.com/pinskylab/OceanAdapt/blob/master/R/add-spp-to-taxonomy.Rmd#L33

Visualization of flagged taxa

```{r, echo=FALSE, out.width = '80%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "taxonomic_flagging", paste0(survey$survey[1],"_taxonomic_flagging.png")))
```

Statistics related to the taxonomic flagging outputs

```{r, echo=FALSE}
df <- read.csv(here::here("standardization_steps", "outputs", "taxonomic_flagging", paste0(survey$survey[1],'_stats.csv')))
knitr::kable(df, col.names = NULL)
```

## 10. Spatio-temporal standardization

### a. Standardization method 1

This standardization method was adapted from https://github.com/zoekitchel/trawl_spatial_turnover/blob/master/data_prep_code/species/explore_NorthSea_trimming.Rmd
\
It was run for hex resolution 7 and 8.

Plot of number of cells x years with overlaid flagging options

```{r, echo=FALSE, out.width = '80%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res7", paste0(survey$survey[1],"_hex_res_7_plot.png")))
```
```{r, echo=FALSE, out.width = '80%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_plot.png")))
```

Map of hauls retained and removed per flagging method and threshold

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res7",  paste0(survey$survey[1],"_hex_res_7_map_per_haul.png")))
```

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_map_per_haul.png")))
```


Map of numbers of years removed per grid cell and flagging method/threshold

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res7", paste0(survey$survey[1],"_hex_res_7_map_per_grid_nyears.png")))
```


```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_map_per_grid_nyears.png")))
```


### b. Standardization method 2

This standardization method was adapted from BioTIME code from https://github.com/Wubing-Xu/Range_size_winners_losers 
   
Map of hauls retained and removed

```{r, echo=FALSE, out.width = '100%'}
knitr::include_graphics(here::here("standardization_steps", "outputs", "trimming_method2", 
                                   paste0(survey$survey[1],"_map_per_haul.png")))
```

### c. Standardization summary

Statistics of hauls removed for each standardization method

```{r, echo=FALSE}
met1_7 <- read.csv(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res7", paste0(survey$survey[1],"_hex_res_7_stats_hauls.csv")))
met1_8 <- read.csv(here::here("standardization_steps", "outputs", "trimming_method1", "hex_res8", paste0(survey$survey[1],"_hex_res_8_stats_hauls.csv")))
met2 <- read.csv(here::here("standardization_steps", "outputs", "trimming_method2", 
                            paste0(survey$survey[1],"_stats_hauls.csv")))
knitr::kable(cbind(met1_7, met1_8[,2:3], met2[,2]), 
             col.names = c("summary", "grid cell 7, 0% threshold", "grid cell 7, 2% threshold", 
                           "grid cell 8, 0% threshold", "grid cell 8, 2% threshold",
                           "method 2 (biotime)")) %>% 
  kable_styling(latex_options = "scale_down",
                position = "float_left")
```


```{r end, echo=FALSE, eval =F, results='hide', include=FALSE}
#THIS CODE REMOVES THE ORIGINAL SURVEY FROM YOUR COMPUTER.
# IT WILL NOT PRINT IN THE FINAL DOCUMENT
file.remove(here("./summary/DATRAS_v3_clean.csv")) # changed for multiple users
```
